Kids First or Marriage First?

### 2019-06-02 main/draft.tex
Many edits, preparing estimation section

\documentclass[12pt,letter]{article}
\usepackage[left=0.8in,right=0.8in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{microtype}
\usepackage{euscript}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{mathrsfs}
\usepackage{tikz}
\newcommand{\hypo}{\mathcal{H}}            
\bibliographystyle{ieeetr}

@@

\begin{abstract}
Unplanned pregnancies often trigger marriage for unmarried couples. I show that, on average, marriages in which a child appeared before or at the year of marriage are characterized with higher divorce probabilities and other outcomes suggesting lower marital quality. This association is robust to different ways of controlling for composition, and it is more pronounced for higher income and more educated couples. I argue that this pattern is consistent with household bargaining models, and the main driver of the result is unequal exposure of spouses to the childcare costs: since being single mother is a bad outcome for an educated female, she is more likely to agree to enter the marriages of lower quality. I estimate a lifecycle model with dynamic bargaining and fertility decisions to show how much of these differences is driven by household bargaining and how unplanned pregnancies creating unexpected childcare costs contribute to observable divorce rates.
\end{abstract}

@@

\section{Empirical Strategy}
The key empirical observation is association between marriage timing and divorce probabilities. The goal is to argue that statistically women who have their kids before they are married have significantly higher chance to be divorced. American Community Survey (ACS) data are suitable for establishing this fact only of particular subsample of women: most importantly, those who had no more than one marriage in their life. To re-confirm the patterns I additionally use Survey of Income and Program Participation (SIPP) data: they have smaller sample size but collect exact questions I care about.

\subsection{Defining Variables}

@@

\subsection{Defining Income}
This part describes how I define income partitions in ACS and SIPP. Shortly, I define income groups based on per hour income of working females, by excluding those who does not work from income considerations. In the appendix \ref{inc-part} I try several alternative partitions, shortly they do not affect the conclusions although they do affect magnitudes in different groups.

To define per-hour income, I restrict attention to working women only. I define working as those who are employed now, report to work more than 25 weeks in the last year, report non-missing labor income above or equal to $\$1000$ per year, report non-missing usual working hours above or equal to $5$ hours per week.

@@

In SIPP 2014, I perform the similar operations. I exclude females who had less than $1000/12$ monthly income or had average weekly hours less than 5 during the last month. Then I divide monthly income on four times average weekly hours to get estimate for per-hour earnings. As before, I residualize logs of these income by regressing them on the same factors, except that instead of year controls (I only use 2014) I use month controls that allow to capture variation in number of work days per month.

\section{Reduced Form Evidence}
\subsection{Raw Comparisons}
The main result I discuss is summarized by Figure \ref{shrs}. Table \ref{diff-raw} presents these results in numbers.

\begin{figure}[h!]
\includegraphics[width=0.9\linewidth]{shrs.eps}
\caption{Share of K$\to$M couples and share of divorced\label{shrs}}
\end{figure}

%First, I compare divorce probabilities in cross-sections of people between $K\to M$ and $M\to K$ groups. See Table \ref{diff-raw} for this.

\begin{table}[p]
\begin{center}
\begin{tabular}{|l|c||c|c||c||c|c|}\cline{2-7}
\multicolumn{1}{c|}{} & \multicolumn{3}{c||}{\small ACS: sample of married once} & \multicolumn{3}{c|}{\small SIPP: sample of  ever married} \\\cline{2-7}
\multicolumn{1}{c|}{}  & \multirow{2}{*}{\small \% K$\to$M}& \multicolumn{2}{c||}{\textbf{\% divorced (now)}} & \multirow{2}{*}{\footnotesize \% K$\to$M }& \multicolumn{2}{c|}{\textbf{\% divorced (ever)}} \\\cline{3-4}\cline{6-7}
\multicolumn{1}{c|}{}  &  &\small if M$\to$K &  \small if K$\to$M & & \small if M$\to$K & \small if K$\to$M \\\hline
\textit{All females, 21--40}  & $23.2\%$ & $10.1\%$ & $18.0\%$ & $22.8\%$ & $24.1\%$ & $34.8\%$ \\
\textit{All females, 35--40} & $18.0\%$ & $11.3\%$ & $23.0\%$ & $18.9\%$ & $28.1\%$ & $45.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 21--40} & $34.2\%$ & $13.9\%$ & $17.2\%$ & $33.4\%$ & $29.2\%$  & $34.2\%$ \\
\textit{College graduates, 21--40} & $9.9\%$ & $5.3\%$ & $14.7\%$ & $10.5\%$ & $12.9\%$ & $27.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} per-hour income for subsample of working women.}\\\hline
\textit{Low 30\% income, 21--40} & $30.8\%$ & $17.6\%$ & $24.1\%$ & $34.6\%$ & $30.5\%$ &  $42.5\%$ \\
\textit{Mid 40\% income, 21--40} & $25.5\%$ & $14.2\%$ & $21.5\%$ & $24.1\%$ & $28.3\%$ & $45.8\%$ \\
\textit{Top 30\% income, 21--40} & $14.1\%$ & $7.7\%$ & $18.1\%$ & $12.4\%$ & $22.9\%$  & $32.8\%$ \\\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$.}\\
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.}\\
\hline
\end{tabular}
\caption{Differences in raw means, ACS and SIPP\label{diff-raw}}
\end{center}
\end{table}

@@

This alone is not enough to draw any causal conclusions: composition of these groups is drastically different. For instance, as suggested by the table, people with lower income are more likely to get divorced, so even conditional on income/education group K$\to$M couples may be poorer with each subgroup. Additionally, many other factors may drive the differences, including different ages of children (people in K$\to$M group by construction have older babies after any given number of years after marriage), regional and racial differences in attitudes to divorce, age composition, possible presence of step-children in K$\to$M couples and so on. I attempt to control for few of these things and argue that they explain these differences only partially, that suggests the presence of causal mechanism that I describe.

\subsubsection{Duration Considerations}
One obvious concern about comparing two groups is the temporal nature of the data: the variable I use captures fact that divorce happened in the past, and the exact timing may be different between the two groups. Due to cumulative nature of the divorce indicator, females who had marriage long time ago are more likely to be divorced even if their divorce chances in any given period are the same. Therefore, I it is useful to look at divorced females conditional on number of years passed from marriage, see the left panel of Figure \ref{pic-aftmar} for this.
\begin{figure}[h!]
\includegraphics[width=0.45\linewidth]{div_by_aftmar.eps}
\includegraphics[width=0.45\linewidth]{aftmar_by_mk.eps}
\caption{Divorced by years after marriage and distribution of years after marriage\label{pic-aftmar}}
\end{figure}

\subsection{Adjusting for Composition}

This part documents large differences in variables that may affect divorce rates between the two groups. After that I try to flexibly capture these factors to see what part of the difference is attributed to the fact of being in K$\to$M group. These exercises use ACS data as they have more observations so the tests are less noisy. Shortly, after adding many controls, observable difference decreases about twice, but still remains sizable and significant.

\subsection{Differences in Observables}
This section captures demographic and economic differences of K$\to$M and M$\to$K groups. Many factors could compute to the observable differences, here I quantify them and this justifies use of controls in the next part. I present these differences for ACS, SIPP data are not very different except smaller sample size. For ACS, since M$\to$K and K$\to$M are defined only for subsample of people who were married once and have children present, I also compare this subsample with more general female population. See Table \ref{diff-contr} for the results.

\subsubsection{Statistical Tests and Controls}

In this section I formally test differences in shares for different subgroups and attempt to control for composition using regressions with many fixed effects.

@@

%The most important controls are age of the eldest child and number of years after marriage. Basically, being in K$\to$M group is defined by their interaction

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|c||c|c|c|c|c|c|}\cline{2-8}
\multicolumn{1}{c|}{} & \multicolumn{7}{c|}{ACS: sample of married once}\\\cline{2-8}
\multicolumn{1}{c|}{} & \small Share of &\multicolumn{2}{c|}{\small Raw difference} & \multicolumn{2}{c|}{\small with controls} & \multicolumn{2}{c|}{\small controls + income} \\ \cline{3-8}
\multicolumn{1}{c|}{} & \small divorced & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) \\  \hline
\textit{All females, 20--50} & 14.3\% &  7.9\% &   (0.1\%) &    5.7\%  &    (0.2\%) &     4.6\%   &   0.2\% \\
\textit{All females, 35--40} &  17.3\% &  11.7\%  &  (0.2\%) &   6.7\% &  (0.3\%) &     6.2\%  &   (0.4\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 20--50} & 18.1\% &   3.3\%  &    (0.2\%) &   8.4\% &   (0.3\%)   &  7.8\%   &  (0.5\%) \\
\textit{College graduates, 20--50} & 8.5\% &   9.4\% &   (0.2\%) &   3.9\% &   (0.3\%) &   3.2\%   &  (0.4\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} working women only.}\\\hline
\textit{Low 30\% income, 20--40} &  21.6\% &  6.5\% &    (0.3\%) &    7.8\% &  (0.5\%) &   7.8\%   &  (0.5\%) \\
\textit{Mid 40\% income, 20--40} &  17.1\% &  7.3\% &   (0.3\%) &    3.6\%   & (0.4\%) &   3.6\% &    (0.4\%) \\
\textit{Top 30\% income, 20--40} & 11.1\% &  10.4\% &  (0.3\%) &    3.5\% &   (0.4\%) & 3.5\% &  (0.4\%) \\
\hline
\end{tabular}
\caption{Differences, controlled for composition, ACS\label{diff-contr}}
\end{center}
\end{table}
\subsection{Additional Comparisons}

Here I show few additional results. First, I present several graphs that suggest relation of divorce chances to child's age. Second, I use alternative variable and consider only people who divorced in the year of survey, as for them duration of marriage and age of the first child is (almost) perfectly observed. The main point of these calculations is to show that divorces seem to be more concentrated around the moment where children get older.

\subsubsection{Effects by Child's Age}
First, I show how share of K$\to$M people varies with age of the eldest child. I focus on subsample children are 18 or younger, as very few people have children above 18 living in the household. This exercise highlights a nuance in the construction of the groups: almost no one in K$\to$M group has children of age less than a year, because that would mean the same year of marriage, first birth and ACS survey. Additionally, since the population is trimmed at the age of 40, having kids around age of 18 means fairly low age of the first birth, that is typically associated with being in M$\to$K group. Left panel of Figure \ref{pics-raw} clearly depicts both issues, and this suggests the importance of using controls. In the specifications above I control both for number of years after marriage and the age of the eldest child, so this should help mitigating these issues.

\subsubsection{Sample of Recently Divorced}
In this exercise I utilize an alternative variable available in ACS --- indicator whether a particular person got a divorce in the year of survey. It can be interpreted as \textit{marginal} divorce probability as opposed to the previous value being \emph{cumulative}. Therefore, for instance, it does not have an upward trend in number of years married, see Figure \ref{pic-now-aftmar}. 
\begin{figure}[h]
\centering
\includegraphics[width=0.45\linewidth]{divnow_by_aftmar.eps}
\caption{Probabilities of getting divorced in the year of survey\label{pic-now-aftmar}}
\end{figure}

\subsection{Intra-Household Allocations}

[I look at earnings, hours and LFP differences] [Grade retention depends on child's age, so need to compare given child's age]

[THIS IS AN OLD VERSION OF TABLE]

\begin{center}
\begin{tabular}{|l||c|c||c|c|}\hline
& \multicolumn{2}{|c||}{\footnotesize \textit{All couples}}  & \multicolumn{2}{|c|}{\footnotesize \textit{Top 40\% of income} } \\\hline
&                       \footnotesize  M$\to$K & \footnotesize  K$\to$M &  \footnotesize  M$\to$K & \footnotesize  K$\to$M \\\hline
\footnotesize \textit{Female in Labor Force}        & $70.3\%$  & $71.4\%$  & $79.3\%$ & $85.3\%$  \\\hline
\footnotesize \textit{Earninigs Ratio}    &  $95.2\%$ & $93.9\%$  &  $95.8\%$ &  $96.4\%$ \\\hline
\footnotesize \textit{Kids' Grade Retention}  & $4.6\%$    & $7.6\%$ &   $3.3\%$   & $5.9\%$ \\\hline
\end{tabular}
\end{center}

@@

\section{Estimation Strategy}
I use techniques of indirect inference, extensively described, for instance, in [...] for estimation of the parameters. Shortly, I select parameters such that features of the real data are similar to the features of the data simulated from model. It is not entirely correct to call the procedure simulated method of moments, as some things that I match are not moments in econometric sense, however it has many similarities with simulated method of moments so I occasionally extend notion of moments to reflect any statistics derived from the data.

\subsection{Overview}
Full list of parameters involves many things:
\begin{itemize}
\item Fundamental parameters of income fluctuation model $(\beta,\sigma,R)$
\item Preferences for child's quality: $(\phi_0,\phi_1,\phi_2,\phi_3)$
\item Parameters for child's quality production function: CES power $\theta_k$ and labor shares by age $(\mu_y,\mu_s)$
\item Unavoidable childcare costs $\underline{q}$
\item Transition probabilities:
\begin{itemize}
\item Probability of pregnancy shock by age: $p^{\text{preg}}_t = \lambda^{\text{preg}}_0 - \lambda^{\text{preg}}_1\cdot t$ (2 parameters)
\item Probability of meeting a partner: $p^{\text{meet}}_t:$ piecewise linear function between $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$ (4 parameters)
\item Probability of re-entering marriage market and corresponding utility ``compensation'' $(p^{\text{out}},S)$. I set $S = 0$ in current calibration (so 1 parameter).
\item Probability that kids grow up (transition from $y$ to $s$): $p_s$ (1 parameter)
\end{itemize}
\item Age profiles of labor income by gender: $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ (6 parameters)
\item Variance of shocks (6 parameters):
\begin{itemize}
\item Individual income shocks $\sigma^{z,f}$, $\sigma^{z,m}$.
\item Couple's surplus shocks $\sigma^{\psi}$
\item Couple's initial surplus $\sigma^{\psi_0}$
\item Assets and productivity of potential partner: $(\sigma^{z,p},\sigma^{a,p})$
\end{itemize}
\end{itemize}

@@

I also externally estimate $\sigma^{z,p}$ as variance of (unpredicted part of) log-difference in spouses earnings, see Appendix \ref{partearn} for this. Based on this calculations I set $\sigma^{z,p} = 0.8$. As I both do not have good assets data and do not aim to replicate assets distribution, I set $\sigma^{a,p} = \sigma^{z,p}$.

@@

In the part with assets, I assume savings can be between 0 and 20 times average labor income in the beginning of life. This bar in pretty low, but it is enough to create reasonable assets variation at the moment most of marriage decisions are made: say, median assets-to-income ratio at 30 is around 1.

\subsection{Estimated Parameters}
Out of 31, 15 parameters are set externally, so I am left with $16$ parameters to estimate: child's parameters $(\phi_0,\phi_1,\phi_2)$, child quality production parameters $(\theta_k,\mu_y,\mu_a)$, variances of initial marriage surplus and its innovations $(\sigma^{\psi_0},\sigma^{\psi})$, meeting probability parameters $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$, pregnancy probability parameters $(\lambda^{\text{preg}}_0,\lambda^{\text{preg}}_1)$, transition out of being single mother parameters $p^{\text{out}}$, $S$. Unfortunately, all of them interact and as the model is highly stylized but involves many choices, they cannot be estimated externally.

>>>>

\documentclass[12pt,letter]{article}
\usepackage[left=0.8in,right=0.8in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{microtype}
\usepackage{euscript}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{mathrsfs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\tikzstyle{block} = [rectangle, draw, rounded corners]
\tikzstyle{line} = [draw, -latex']
\newcommand{\hypo}{\mathcal{H}}            
\bibliographystyle{ieeetr}

@@

\begin{abstract}
Unplanned pregnancies often trigger marriage for unmarried couples. I document two robust empirical facts. First, women who have their first child before or at the year of marriage are statistically more likely to be divorced in the future than those who marry before they have a child. Second, this difference is more pronounced for women with higher income and educational attainment, although it is less common for them to have their kids before marriage. I estimate a lifecycle model with dynamic bargaining, marriage selection and unplanned pregnancy risks. I show that the first fact can be attributed to household bargaining friction, as women agree to enter lower quality marriages to avoid being a single mother, and  the second fact is consistent with childcare requiring both time and money, as this suggests that the costs of being a single mother are higher for more productive women. The model suggests large causal effect of unplanned pregnancies on observable divorce rates. 
\end{abstract}

@@

\section{Empirical Facts}

This section presents key empirical facts and several exercises on testing their validity. In summary:
\begin{enumerate}
\item Women who had their first child before their first marriage (K$\to$M) are more likely to be divorced than those who had their first child in their first marriage (M$\to$K).
\item Relative difference in the share of divorced between K$\to$M and M$\to$K is increasing with income and education
\item These results are statistically significant and robust to various methods of controlling for observable composition of the groups
\end{enumerate}

To establish these results, I mainly use American Community Survey (ACS), that has large sample sizes and many relevant observable characteristics of households. Its only limitation is that it does not recover marriage and fertility history perfectly. Because of this, I re-confirm the results with the data from Survey of Income and Program Participation (SIPP 2014), that has fewer observations but more detailed marriage and fertility history.

I begin with describing empirical strategy, then Section \ref{keycomp} presents the key results for ACS and SIPP, the following sections present several validity checks.

\subsection{Defining Groups}

@@

\subsection{Income Partition}
I define income groups based on per hour income of working females, by excluding those who does not work or work too little from income considerations. Precisely, I define working as those who are employed now, report to work more than 25 weeks in the last year, report non-missing labor income above or equal to $\$1000$ per year, report non-missing usual working hours above or equal to $5$ hours per week.

@@

In SIPP 2014, I perform the similar operations. I exclude females who had less than $1000/12$ monthly income or had average weekly hours less than 5 during the last month. Then I divide monthly income on four times average weekly hours to get estimate for per-hour earnings. As before, I residualize logs of these income by regressing them on the same factors, except that instead of year controls (I only use 2014) I use month controls.

\subsection{Raw Comparisons\label{keycomp}}
The main results I discuss is summarized by Figure \ref{shrs} for ACS and Figure \ref{shrs-sipp} for SIPP. Table \ref{diff-raw} in presents these results in numbers.

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\linewidth]{shrs.eps}
\begin{tabular}{|p{0.65\linewidth}|}
\hline
\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$\\
\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.\\\hline
\end{tabular}
\caption{Share of K$\to$M couples and share of divorced, ACS\label{shrs}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\linewidth]{shrs-sipp.eps}
\caption{Share of K$\to$M couples and share of ever divorced, SIPP\label{shrs-sipp}}
\end{figure}

\begin{table}[p]
\begin{center}
\begin{tabular}{|l|c||c|c||c||c|c|}\cline{2-7}
\multicolumn{1}{c|}{} & \multicolumn{3}{c||}{\small ACS: sample of married once} & \multicolumn{3}{c|}{\small SIPP: sample of  ever married} \\\cline{2-7}
\multicolumn{1}{c|}{}  & \multirow{2}{*}{\small \% K$\to$M}& \multicolumn{2}{c||}{\textbf{\% divorced (now)}} & \multirow{2}{*}{\footnotesize \% K$\to$M }& \multicolumn{2}{c|}{\textbf{\% divorced (ever)}} \\\cline{3-4}\cline{6-7}
\multicolumn{1}{c|}{}  &  &\small if M$\to$K &  \small if K$\to$M & & \small if M$\to$K & \small if K$\to$M \\\hline
\textit{All females, 21--40}  & $23.2\%$ & $10.1\%$ & $18.0\%$ & $22.8\%$ & $24.1\%$ & $34.8\%$ \\
\textit{All females, 35--40} & $18.0\%$ & $11.3\%$ & $23.0\%$ & $18.9\%$ & $28.1\%$ & $45.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 21--40} & $34.2\%$ & $13.9\%$ & $17.2\%$ & $33.4\%$ & $29.2\%$  & $34.2\%$ \\
\textit{College graduates, 21--40} & $9.9\%$ & $5.3\%$ & $14.7\%$ & $10.5\%$ & $12.9\%$ & $27.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} per-hour income for subsample of working women.}\\\hline
\textit{Low 30\% income, 21--40} & $30.8\%$ & $17.6\%$ & $24.1\%$ & $34.6\%$ & $30.5\%$ &  $42.5\%$ \\
\textit{Mid 40\% income, 21--40} & $25.5\%$ & $14.2\%$ & $21.5\%$ & $24.1\%$ & $28.3\%$ & $45.8\%$ \\
\textit{Top 30\% income, 21--40} & $14.1\%$ & $7.7\%$ & $18.1\%$ & $12.4\%$ & $22.9\%$  & $32.8\%$ \\\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$.}\\
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.}\\
\hline
\end{tabular}
\caption{Differences in raw means, ACS and SIPP (repeats Figures \ref{shrs} and \ref{shrs-sipp}) \label{diff-raw}}
\end{center}
\end{table}

@@

This alone is not enough to draw strong conclusions: composition of these groups is drastically different. For instance, as suggested by the table, people with lower income are more likely to get divorced, so even conditional on income/education group K$\to$M couples may be poorer with each subgroup. Additionally, many other factors may drive the differences, including different ages of children (people in K$\to$M group by construction have older babies after any given number of years after marriage), regional and racial differences in attitudes to divorce, age composition, possible presence of step-children in K$\to$M couples and so on. I attempt to control for few of these things and argue that they explain these differences only partially, that suggests the presence of causal mechanism that I describe.

\subsection{Adjusting for Composition}
As people in K$\to$M and M$\to$K groups obviously have different characteristics, as in particular suggested by Figure \ref{keycomp}, I use regression techniques to see what part of difference is attributed to this composition. Appendix \ref{comp-diff-appendix} quantifies these differences, shortly, people in K$\to$M group are younger, less educated, have kids earlier, marry slightly earlier and have more unequal marriages. In this part I control for difference in observable characteristics to see what part of the difference is attributed to the fact of being in K$\to$M group. Shortly, controlling for observables mitigate some part of the difference but it still remains sizable and significant.

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|c||c|c|c|c|c|c|}\cline{2-8}
\multicolumn{1}{c|}{} & \multicolumn{7}{c|}{ACS: sample of married once}\\\cline{2-8}
\multicolumn{1}{c|}{} & \small Share of &\multicolumn{2}{c|}{\small Raw difference} & \multicolumn{2}{c|}{\small with controls} & \multicolumn{2}{c|}{\small controls + income} \\ \cline{3-8}
\multicolumn{1}{c|}{} & \small divorced & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) \\  \hline
\textit{All females, 20--50} & 14.3\% &  7.9\% &   (0.1\%) &    5.7\%  &    (0.2\%) &     4.6\%   &   0.2\% \\
\textit{All females, 35--40} &  17.3\% &  11.7\%  &  (0.2\%) &   6.7\% &  (0.3\%) &     6.2\%  &   (0.4\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 20--50} & 18.1\% &   3.3\%  &    (0.2\%) &   8.4\% &   (0.3\%)   &  7.8\%   &  (0.5\%) \\
\textit{College graduates, 20--50} & 8.5\% &   9.4\% &   (0.2\%) &   3.9\% &   (0.3\%) &   3.2\%   &  (0.4\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} working women only.}\\\hline
\textit{Low 30\% income, 20--40} &  21.6\% &  6.5\% &    (0.3\%) &    7.8\% &  (0.5\%) &   7.8\%   &  (0.5\%) \\
\textit{Mid 40\% income, 20--40} &  17.1\% &  7.3\% &   (0.3\%) &    3.6\%   & (0.4\%) &   3.6\% &    (0.4\%) \\
\textit{Top 30\% income, 20--40} & 11.1\% &  10.4\% &  (0.3\%) &    3.5\% &   (0.4\%) & 3.5\% &  (0.4\%) \\
\hline
\end{tabular}
\caption{Differences, controlled for composition, ACS\label{diff-contr}}
\end{center}
\end{table}

\subsection{Additional Comparisons}

Appendix \ref{extra-comparisons} shows few additional graphs and exercise. I summarize their results here:
\begin{enumerate}
\item In the sample I use M$\to$K women are married for longer at the moment of survey\footnote{This does not contradict the fact that K$\to$M couples marry slightly earlier as I use the sample of married women.} (so they had \emph{more} years to break up), comparing marriages of the same duration only exacerbates the difference.
\item Children of K$\to$M women are on average older (as they are born in earlier ages). Divergence between K$\to$M and M$\to$K observable divorce probabilities happens around ages 5--10. After controlling for composition, however, I show that K$\to$M couples divorce more often even conditional on child's age.
\item To further address concerns about duration, I recompute the numbers using getting the divorce at the year of survey, corresponding to marginal divorce probability as opposed to cumulative divorce probability (being divorced at by the moment survey). The magnitudes and patterns are very similar, though the numbers are harder to interpret. This suggests that 
\item I show that the difference is not likely to be driven by presence of stepchildren in K$\to$M families. In fact, as I restrict $\Delta T \geq -5$, share of stepchildren is relatively small, and surviving couples with stepchildren do not differ in observable marriage duration with couples without stepchildren.
\end{enumerate}

\subsection{Intra-Household Allocations [TBD]}
I document few additional facts describing the difference between K$\to$M and M$\to$K besides being divorced. Following Lafortune et al, 2018, I construct measure of children's grade retention based on ACS (that is an indicator if any school-age child in the household is in grade that is lower than prescribed by her/his age). I show this together with few other intra-houeshold characteristics: measure of inequality and labor force participation.

See Table \ref{intra} for the numbers. The exercise suggests that children on K$\to$M women repeat grades significantly more often, and this happens even in higher-income households. [This table is based on old sample so it is slightly inconsistent with what I do above, I will redo this in one of the future iterations.]

\begin{table}
\centering
\begin{tabular}{|l||c|c||c|c|}\hline
& \multicolumn{2}{|c||}{\footnotesize \textit{All couples}}  & \multicolumn{2}{|c|}{\footnotesize \textit{Top 40\% of income} } \\\hline
&                       \footnotesize  M$\to$K & \footnotesize  K$\to$M &  \footnotesize  M$\to$K & \footnotesize  K$\to$M \\\hline
\footnotesize \textit{Female in Labor Force}        & $70.3\%$  & $71.4\%$  & $79.3\%$ & $85.3\%$  \\\hline
\footnotesize \textit{Earninigs Ratio}    &  $95.2\%$ & $93.9\%$  &  $95.8\%$ &  $96.4\%$ \\\hline
\footnotesize \textit{Kids' Grade Retention}  & $4.6\%$    & $7.6\%$ &   $3.3\%$   & $5.9\%$ \\\hline
\end{tabular}
\caption{Child's outcomes and intra-household allocations [to be fixed later]\label{intra}}
\end{table}

%\section{Mechanisms}
%This part describes the mechanism I am trying to capture in the model that is presented in the following section. Suppose single agents with characteristics $\omega^m$ and $\omega^f$ can meet each other and form a couple. On top of individual characteristics, let couples be characterized by idiosyncratic utility gain $\psi$ from being together, that can be interpreted as a love shock. Let also there exists a variable $\theta$ that describes terms of marriage (that can be interpreted as decision weight of wife in a collective household). Let for $i \in \{f,m\}$ functions $V^{i,c}$ and $V^{i,s}$ denote values of being in a couple and being single, respectively. People agree to marry if there exist marriage terms $\theta$ such that gains of both spouses from marriage is positive, that is
%\[\Delta^f = V^{f,c}(\omega^f,\omega^m,\theta,\psi) - V^{f,s}(\omega^f) \geq 0, \ \ \Delta^m = V^{m,c}(\omega^f,\omega^m,\theta,\psi) -  V^{m,s}(\omega^m) \geq 0.\]
%
%Assume that utilities of potential spouses increase in $\psi$. Assume also that wife benefits from higher $\theta$ and husband benefits from lower $\theta$. The following illustration captures the intuition for the model:

@@

\section{Quantification}
I use techniques of indirect inference to fit the model to the data. Namely, given the parameter values I solve for decision rules and simulate $5{,}000$ agents for the first 20 periods of their life corresponding to ages of (21--40 used in the data part). During estimation procedure I look for the parameter values such that observable differences between between the simulated and actual data, measured by the targets I describe below, is minimal.

%\subsection{Overview}
%Full list of parameters involves many things:
%\begin{itemize}
%\item Fundamental parameters of income fluctuation model $(\beta,\sigma,R)$
%\item Preferences for child's quality: $(\phi_0,\phi_1,\phi_2,\phi_3)$
%\item Parameters for child's quality production function: CES power $\theta_k$ and labor shares by age $(\mu_y,\mu_s)$
%\item Unavoidable childcare costs $\underline{q}$
%\item Transition probabilities:
%\begin{itemize}
%\item Probability of pregnancy shock by age: $p^{\text{preg}}_t = \lambda^{\text{preg}}_0 - \lambda^{\text{preg}}_1\cdot t$ (2 parameters)
%\item Probability of meeting a partner: $p^{\text{meet}}_t:$ piecewise linear function between $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$ (4 parameters)
%\item Probability of re-entering marriage market and corresponding utility ``compensation'' $(p^{\text{out}},S)$. I set $S = 0$ in current calibration (so 1 parameter).
%\item Probability that kids grow up (transition from $y$ to $s$): $p_s$ (1 parameter)
%\end{itemize}
%\item Age profiles of labor income by gender: $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ (6 parameters)
%\item Variance of shocks (6 parameters):
%\begin{itemize}
%\item Individual income shocks $\sigma^{z,f}$, $\sigma^{z,m}$.
%\item Couple's surplus shocks $\sigma^{\psi}$
%\item Couple's initial surplus $\sigma^{\psi_0}$
%\item Assets and productivity of potential partner: $(\sigma^{z,p},\sigma^{a,p})$
%\end{itemize}
%\end{itemize}

@@

I also externally estimate $\sigma^{z,p}$ as variance of (unpredicted part of) log-difference in spouses earnings, see Appendix \ref{partearn} for this. Based on this calculations I set $\sigma^{z,p} = 0.6$ (this reproduces $\sigma = 0.7$ in the simulated data because income shocks are uncorrelated in the model as opposed to the data). As I both do not have good assets data and do not aim to replicate assets distribution, I set $\sigma^{a,p} = \sigma^{z,p}$.

@@

In the part with assets, I assume savings can be between 0 and 20 times average labor income in the beginning of life. This bar in pretty low, but it is enough to create reasonable assets variation at the moment most of marriage decisions are made, creating median assets-to-income ratio at 30 being slightly above 1.

%\subsection{Estimated Parameters}
%Out of 31, 15 parameters are set externally, so I am left with $16$ parameters to estimate: child's parameters $(\phi_0,\phi_1,\phi_2)$, child quality production parameters $(\theta_k,\mu_y,\mu_a)$, variances of initial marriage surplus and its innovations $(\sigma^{\psi_0},\sigma^{\psi})$, meeting probability parameters $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$, pregnancy probability parameters $(\lambda^{\text{preg}}_0,\lambda^{\text{preg}}_1)$, transition out of being single mother parameters $p^{\text{out}}$, $S$. Unfortunately, all of them interact and as the model is highly stylized but involves many choices, they cannot be estimated externally.
